## text

This part deals with all text-related things in the format. Both in terms of typography and some rules about general writing

### Table showing all font sizes, families etc.

put the table here.

| level             | font      | size | styling | add? |
|-------------------|-----------|------|---------|------|
| H1 - report title | Open sans |      |         |      |
| H2                | Open sans |      |         |      |
| H3                | Open sans |      |         |      |
| H4                | Open sans |      |         |      |
| regular text      | Open sans | 12pt |         |      |
| caption           | Open sans |      |         |      |
| axis titles       | Open sans |      |         |      |
| axis text         | Open sans |      |         |      |
| plot text         | Open sans |      |         |      |

Are there any additional rules about text?

Probably something like:

-   avoid very long titles for reports

-   avoid very long headings

-   make sure text does not overlap -\> `{ggrepel}`\
    This package has a function `geom_text_repel()` which tries to adjust labels so that they don't overlap. For example this plot has an issue with overlap:

    ```{r}

    ```

    We can deal with this using `geom_text_repel()`. Additional arguments like `force` control how much the labels should be pushed from their original positions.

    ```{r}

    ```

-   ensure proper contrast for text -\> `{ggfittext}`

    For example the plot below does not have decent contrast when all text is in black:\

    ```{r}

    ```

    We can fix that using `{ggfittext}` and `geom_fit_text()`

    ```{r}

    ```

General rules about writing should go here?

-   Be accessible... Don't write about logits. Opt for Average Marginal Effects in percentages instead. If possible try to get a sense of scale and some interpretation for the effects (maybe something like probability of superiority over cohen's d etc.)

-   But remain scientific. Be precise, don't overstate - this is scientific writing.
