## text

This part deals with all text-related things in the format. Both in terms of typography and some rules about general writing. The font used throughout the report is Jost. The text hierarchy is defined below:

| level | font | size | styling |
|-------------------|------------------|------------------|------------------|
| [H1 - report title]{style="font-size: 2.0em;"} | Open sans | 2em |  |
| [H2 - main points]{style="font-size: 1.68em;"} | Open sans | 1.68em |  |
| [H3 - section title]{style="font-size: 1.41em;"} | Open sans | 1.41em |  |
| [H4 - subsection]{style="font-size: 1.189em; font-weight bold;"} | Open sans | 1.1892em |  |
| regular text | Open sans | 12pt |  |
| [caption]{style="font-size: 8pt;"} | Open sans | 8pt | grey, name and number bold |
| [axis titles]{style="font-size: 8.5pt;"} | Open sans | 8.5pt |  |
| [axis text]{style="font-size: 8pt;"} | Open sans | 8pt |  |
| [plot text]{style="font-size: 10pt;"} | Open sans | 10pt |  |

## Some additional rules and tips

-   avoid very long titles for reports

-   avoid very long section titles

-   make sure text does not overlap on plots -\>\
    `{ggrepel}` package has a function `geom_text_repel()` which tries to adjust labels so that they don't overlap. For example this plot has an issue with overlap:

    ```{r warning=FALSE, message=FALSE}
    library(dplyr)
    library(ggplot2)
    graph_df <- data.frame(
      target = c(rep("group 1", 6), rep("group 2", 6)),
      year =  c(rep(2017, 3), rep(2021, 3),rep(2017, 3), rep(2021, 3)),
      att = factor(c(rep(c("ok", "meh", "not ok"), 4)), ordered = T, levels = c("ok", "meh", "not ok")),
      value = c(.0714, .3121, .6165, .034, .2021, .7639, .0189, .1859, .7952, .0103, .0964, .8933)
    )

    graph_df %>%
      ggplot(aes(x = target, y = value, fill = att)) +
      geom_col(width = .6) +
      geom_text(aes(label = paste0(value*100, "%")),direction = "y", position = position_stack(vjust = .5)) +
      scale_fill_manual(values= c("ok"= "steelblue",
                                  "meh" = "grey80",
                                  "not ok" = "tomato4")) +
      facet_wrap(~year) +
      labs(x = NULL, y = NULL) +
      theme_minimal() +
      theme(legend.title = element_blank())
    ```

    We can deal with this using `geom_text_repel()`. Additional arguments like `force` control how much the labels should be pushed from their original positions.

    ```{r}
    library(ggrepel)

    graph_df %>%
      ggplot(aes(x = target, y = value, fill = att)) +
      geom_col(width = .6) +
      geom_text_repel(aes(label = paste0(value*100, "%")),direction = "y", position = position_stack(vjust = .5)) +
      scale_fill_manual(values= c("ok"= "steelblue",
                                  "meh" = "grey80",
                                  "not ok" = "tomato4")) +
      facet_wrap(~year) +
      labs(x = NULL, y = NULL) +
      theme_minimal() +
      theme(legend.title = element_blank())
    ```

-   ensure proper contrast for text:

    For example the plot below does not have decent contrast when all text is in black:

    ```{r}
    b_df <- data.frame(
        a = sample(letters[1:5], 1e3, replace = T),
        b = sample(letters[1:3], 1e3, replace = T)
    ) %>%
    count(a,b)

    b_df %>%
    ggplot(aes(x = a, y = n, fill = b)) +
    geom_col(position = "fill") +
    geom_text(aes(label = n),direction = "y", position = "fill", vjust = 2) +
    scale_fill_manual(values = c("grey80", "firebrick4", "mistyrose")) +
    theme_minimal() +
    theme(legend.title = element_blank(),
         legend.position = "top")
    ```

    We can fix that using `{ggfittext}` and `geom_fit_text():`

    ```{r warning = FALSE}
    library(ggfittext)

    b_df %>%
    ggplot(aes(x = a, y = n, fill = b)) +
    geom_col(position = "fill") +
    geom_fit_text(aes(label = n, fill = b), position = "fill", vjust = 2, contrast = TRUE, show_guide = FALSE) +
    scale_fill_manual(values = c("grey80", "firebrick4", "mistyrose")) +
    theme_minimal() +
    theme(legend.title = element_blank(),
         legend.position = "top")
    ```

-   Be accessible... Don't write about logits. Opt for Average Marginal Effects in percentages instead. If possible try to get a sense of scale and some interpretation for the effects (maybe something like probability of superiority over cohen's d etc.)

-   But remain scientific. Be precise, don't overstate - this is scientific writing.